100 Years of Social Work Research: A Data Science Perspective

### 2014-12-30 Analysis.Rmd
Updated notes.

The search results were exported in the _Generic bibliographic format_.  This format is a text (*.txt) file that is processed with the `BibWrangleR` functions.

@@

This section processes raw data.  This section of code is executed only one time to transform the raw data into an analyzable format.  When new data are obtained for this study (i.e., updated search results), this section should be re-run by changing `echo=FALSE` to `echo=TRUE` in the knitr markdown argument.

```{r initialization, warning=FALSE, message=FALSE, eval=FALSE}
# Clear workspace
rm(list=ls())

# Read BWR functions
source("/Users/beperron/Git/BibWrangleR/functions/piWrangleR.R")
source("/Users/beperron/Git/BibWrangleR/functions/packages.R")

@@

```{r message=FALSE, warning=FALSE, comment=NA}
rm(list=ls())
load("piArticles.R")
library(dplyr)
library(ggplot2)
library(gridExtra)

# Dimensions of the data file (Rows X Columns)
dim(pi.df)

#Variables of the data file
names(pi.df)

@@

# Overall number and name of unique journal titles?

```{r}
unique.titles <- filter(pi.df, attributes == "SO")

#Number of unique titles
length(unique(unique.titles$record))

#Unique titles
unique(unique.titles$record)
```

# Number of unique journal titles by year

```{r message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4}
journals.year <- tbl_df(pi.df)
year <- journals.year %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)

@@

# What journals published the most number of articles

```{r}

n.so.yr <- filter(pi.df, attributes == "SO" | attributes == "YR")
n.so <- filter(pi.df, attributes == "SO") %>% mutate(title = record) %>% select(-attributes, -record)
n.yr <- filter(pi.df, attributes == "YR") %>% mutate(year = record ) %>% select(-attributes, -record)
n.so.yr <- left_join(n.so, n.yr) %>%
    group_by(title) %>%
    summarise(
        first = min(year),
        last = max(year),
        n.to.date = n()
        ) %>%
    arrange(desc(n.to.date))

#10 highest number of publications
head(n.so.yr, 10)
```

# What is the lifespan of journals?  
```{r}
#10 longest running journals 
longest.running <- n.so.yr %>%
       mutate(last = as.numeric(last), first = as.numeric(first),year.diff = last - first) %>%
       arrange(desc(year.diff)) %>%
       select(title, first, last, year.diff)
head(longest.running, 10)

#10 shortest running journals
shortest.running <- longest.running %>% arrange(year.diff, first, last)
head(shortest.running, 20)
```
#Number of articles per year

```{r}
n.articles.year <- filter(pi.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

@@

head(df, 10)
tail(df,10)

plot.article.count <- ggplot(df, aes(as.factor(years), y= year.count, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("year") + 
    ylab("count") + 
    ggtitle("Number of Studies by Year") +
    scale_x_discrete(breaks=c(seq(1914, 2014, 10))) +
    scale_y_continuous(breaks = c(seq(0, 2000, 250)))

df$years <- as.numeric(as.character(df$years))
plot.article.cumulative <- ggplot(df, aes(x = years, y = cumsum(year.count))) + 
    geom_line() + 
    theme(axis.text.x = element_text(angle=45, hjust=1)) +
    scale_x_continuous(breaks=pretty(df$years)) +
    xlab("year") +  
    ylab("count") +
    scale_x_continuous(breaks = c(seq(1914,2014,10))) +
    scale_y_continuous(breaks = c(seq(0, 25000, 2500))) + 
    ggtitle("Cumulative Frequency")

grid.arrange(plot.article.count, plot.article.cumulative, ncol=2)
```

# Topic areas (by Subject Terms)

df.2 <- filter(pi.df, attributes == "SU")

subject.terms <- stringr::str_split(df.2$record, pattern = ";")
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))
subject.terms.total <- length(unlist(lapply(subject.terms, function(x) gsub(" ", "", x))))
subject.terms.unique <- length(unique(subject.terms))

@@

@@

## Most Frequent Subject Terms Over Time (by 5 years)

@@

# Author defined keywords

kp.df <- filter(pi.df, attributes == "KP")

subject.terms <- stringr::str_split(kp.df$record, pattern = ";")
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))
subject.terms.total <- length(unlist(lapply(subject.terms, function(x) gsub(" ", "", x))))
subject.terms.unique <- length(unique(subject.terms))

@@

>>>>

The search results were exported in a _generic bibliographic format_, which is an unstructured text (*.txt) file.The text file was processed using the `BibWrangleR` function created by the first author.

@@

This section processes raw data.  This section of code is executed only one time to transform raw text data into an analyzable format.  When new data are obtained for this study (i.e., updated search results), this section should be re-run by changing `echo=FALSE` to `echo=TRUE` in the knitr markdown argument.

```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=FALSE}
# Clear workspace
rm(list=ls())

# Read BWR functions for Mac OS
source("/Users/beperron/Git/BibWrangleR/functions/piWrangleR.R")
source("/Users/beperron/Git/BibWrangleR/functions/packages.R")

@@

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA}
rm(list=ls())
load("piArticles.R")
library(dplyr)
library(ggplot2)
library(gridExtra)

# Inspect dimensions of the data file (Rows X Columns)
dim(pi.df)

# Inspect variable names of the data file
names(pi.df)

@@

## What is the overall number and names of journal titles?

```{r unique_titles}
unique.titles <- filter(pi.df, attributes == "SO")

# Number of unique titles
length(unique(unique.titles$record))

# Unique titles
unique(unique.titles$record)
```

## Number of unique journal titles by year

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4}
journals.year <- tbl_df(pi.df)

year <- journals.year %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)

@@

## What journals published the most number of articles

```{r journal_output}

n.so.yr <- filter(pi.df, attributes == "SO" | attributes == "YR")

n.so <- filter(pi.df, attributes == "SO") %>% mutate(title = record) %>% 
        select(-attributes, -record)

n.yr <- filter(pi.df, attributes == "YR") %>% mutate(year = record ) %>% 
        select(-attributes, -record)

n.so.yr <- left_join(n.so, n.yr) %>%
    group_by(title) %>%
    summarise(first = min(year), last = max(year), n.to.date = n()) %>%
    arrange(desc(n.to.date))

# 10 highest number of publications
head(n.so.yr, 10)
```

## What is the lifespan of journals?

```{r journal_lifespan}
#10 longest running journals 
longest.running <- n.so.yr %>%
       mutate(last = as.numeric(last), first = as.numeric(first), 
              year.diff = last - first) %>%
       arrange(desc(year.diff)) %>%
       select(title, first, last, year.diff)

head(longest.running, 10)

#10 shortest running journals
shortest.running <- longest.running %>% arrange(year.diff, first, last)
head(shortest.running, 20)
```

## What is the number of articles published per year

```{r articles_per_year}
n.articles.year <- filter(pi.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

@@

plot.article.count <- ggplot(df, aes(as.factor(years), 
                    y = year.count, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("year") + 
    ylab("count") + 
    ggtitle("Number of Studies by Year") +
    scale_x_discrete(breaks=c(seq(1914, 2014, 10))) +
    scale_y_continuous(breaks = c(seq(0, 2000, 250)))

df$years <- as.numeric(as.character(df$years))

plot.article.cumulative <- ggplot(df, aes(x = years, y = cumsum(year.count))) + 
    geom_line() + 
    theme(axis.text.x = element_text(angle=45, hjust=1)) +
    scale_x_continuous(breaks=pretty(df$years)) +
    xlab("year") +  
    ylab("count") +
    scale_x_continuous(breaks = c(seq(1914,2014,10))) +
    scale_y_continuous(breaks = c(seq(0, 25000, 2500))) + 
    ggtitle("Cumulative Frequency")

grid.arrange(plot.article.count, plot.article.cumulative, ncol=2)

# Print most recent ten years
head(df, 10)
```

# What are the topic areas (by Subject Terms)?

```{r subject_terms_overall, comment=NA, warning=FALSE, message=FALSE}

su.df <- filter(pi.df, attributes == "SU")

subject.terms <- stringr::str_split(su.df$record, pattern = ";")
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))

subject.terms.total <- length(unlist(lapply(subject.terms, 
                        function(x) gsub(" ", "", x))))

subject.terms.unique <- length(unique(subject.terms))

@@

most.frequent <- arrange(most.frequent, desc(Freq))

# Print summary statistics in list form
subject.terms.l

# Print 25 most commmon terms
head(most.frequent, 25)

@@

## What are the topic areas over time (by Subject terms)?

```{r subject_terms_overtime}
decade <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

@@

# What are topic areas over time (by author specified keywords)?

```{r comment=NA, warning=FALSE, message=FALSE}
kp.df <- filter(pi.df, attributes == "KP")

subject.terms <- stringr::str_split(kp.df$record, pattern = ";")
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))
subject.terms.total <- length(unlist(lapply(subject.terms, 
                        function(x) gsub(" ", "", x))))

subject.terms.unique <- length(unique(subject.terms))

@@

most.frequent <- arrange(most.frequent, desc(Freq))

# Print summary statistics
print(subject.terms.l)

# Print 25 most frequent
head(most.frequent, 25)
```
